Species motif participation provides unique information about species risk of extinction

### 2023-06-04 manuscript/JAE_reply.tex
SI ready to submit

    [[Some nice closing?]]

@@

            \textbf{R:} Yes, a binomial error distribution would be more appropriate. We have corrected the error and updated our results accordingly and thank the Reviewer for pointing this out. Our results were not strongly affected by this change in statistical methodology, except that the relationship between network size and average consumer persistence was no longer significant.

@@

                \textbf{R:} If the Reviewer is thinking of EcoPath with EcoSim, that type of model only works with quite a small food web with very well-understood interaction strengths and initial population sizes. We are not aware of any dynamical model that is fast over a large food web - certainly this was not the case when some of the authors conducted dynamical simulations of webs with 50-100 species. [[Anyone have a better answer? From the focus on major energy flows I think it's EcoSim they're thinking of...]]

@@

                \textbf{R:} Previous research has shown that omnivory can either be stabilising or destabilising, depending on the precise disturbance, strengths of interactions, etc. Our results, which deal specifically with persistence after bottom-up disturbance, add to that picture of complexity without contradicting earlier findings. Note that we do not say that omnivory is \emph{never} beneficial - only that in our networks, in response to our focal disturbance, is is generally detrimental. We also note that our results show apparent competition as associated with less persistence when disturbance is weak, which is compatible with the possibility that the Reviewer mentions above (though we do not specifically model prey switching, which seems to be what the Reviewer refers to).  In the revision we have expanded our discussion on how the strength and direction of motif participation-persistence relationships vary network-to-network (especially in networks with different connectance). We hope that this addition provides some of the nuance the Reviewer missed from the previous revision.

@@

                None of this is to discount the reality of top-down extinctions or their interest from a scientific standpoint. However, since the Bayesian network framework does not include top-down effects any comment we might make about how top-down effects might affect our results would necessarily be pure speculation. Instead, we have added a reference to another paper exploring the relationship between motif roles and persistence in a dynamical context. That paper, due to the computational limits inherent in the dynamical approach, uses only a single-species disturbance. It does, however, include top-down as well as bottom-up effects and may provide the information the Reviewer is seeking.

>>>>

    Thank you for your time and consideration of our manuscript.

@@

            \textbf{R:} Yes, we agree with the Reviewer that a binomial error distribution would be more appropriate. We have corrected the error and updated our results accordingly and thank the Reviewer for pointing this out. Our results were not strongly affected by this change in statistical methodology, except that the relationship between network size and average consumer persistence was no longer significant.

@@

                \textbf{R:} We do think that our statement here is valid. Topological models for simulating species extinctions are only taking the food web structure in to account, and a secondary extinction will occur when a consumer lost all its resources. So, in comparison to topological models simulating the biomass (or similar) dynamics between many species will require both more parameters and simulation time. If we compare to for example the ATN-model requires parameters describing each species body mass and several parameters for defining for example attack rates and handling times. Similarly, models like EcoPath with EcoSim, is limited to smaller food webs with well-understood interaction strengths and initial population sizes and as such require more parameters which even in simple cases are more complex than topological models. However, these more complex models can model more complex mechanisms than what topological models could.

                %We are not aware of any dynamical model that is fast over a large food web - certainly this was not the case when some of the authors conducted dynamical simulations of webs with 50-100 species. [[Anyone have a better answer? From the focus on major energy flows I think it's EcoSim they're thinking of...]]

@@

                \textbf{R:} Previous research has shown that omnivory can either be stabilising or destabilising, depending on the precise disturbance, strengths of interactions, etc. Our results, which deal specifically with persistence after bottom-up disturbance, add to that picture of complexity without contradicting earlier findings. Note that we do not say that omnivory is never beneficial - only that in our networks, in response to our focal disturbance, is is generally detrimental. We also note that our results show apparent competition as associated with less persistence when disturbance is weak, which is compatible with the possibility that the Reviewer mentions above (though we do not specifically model prey switching, which seems to be what the Reviewer refers to).  In the revision, we have expanded our discussion on how the strength and direction of motif participation-persistence relationships vary network-to-network (especially in networks with different connectance). We hope that this addition provides some of the nuance the Reviewer missed from the previous revision.

@@

                None of this is to discount the reality of top-down extinctions or their interest from a scientific standpoint. However, since the Bayesian network framework does not include top-down effects any comment we might make about how top-down effects might affect our results would necessarily be pure speculation. Instead, we have added a reference to another paper (quoted below) exploring the relationship between motif roles and persistence in a dynamical context. That paper, due to the computational limits inherent in the dynamical approach, uses only a single-species disturbance. It does, however, include top-down as well as bottom-up effects and may provide the information the Reviewer is seeking.

                \begin{quotation}

                    There are limitations to the approach presented here.
                    Most importantly, the Bayesian network framework operates on a strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
                    Here our focus was to analyze bottom-up affects following disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
                    If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
                    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2021_inprep}.
                    In particular, species which frequently appeared in the prey position of the direct competition motif tended to have shorter persistence times after disturbance in that study - possibly due to top-down effects from the pair of predators involved.
                    However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.

                \end{quotation}

### 2023-06-04 manuscript/reorder_main_text.tex
SI ready to submit

    \indent{1. Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.} \\
    2. A species' position in a food web (e.g., its trophic level or number of prey) is known to affect its ability to persist following disturbance.
    3. We therefore expect that a species' participation in three-species motifs --meso-scale structures which provide information on a species' direct and indirect interactions-- will also be related to its probability of persistence.\\
    4. We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.\\
    5. However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.\\
    6. Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied.

@@

    The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional (secondary) extinctions. 
    A species' risk of secondary extinction depends on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
    For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{Allesina2012,binzer2011susceptibility, Eklof2006}.
    These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources if a prey species is lost.
    Similarly, a species' participation in \emph{meso-scale motifs} (which provide a more nuanced description of a species' place in the food web) are related to its persistence~\citep{Cirtwill2015a,Cirtwill2018FoodWebs,Cirtwill2022Oikos}.

    Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
    Some motifs are over-represented in stable food webs, suggesting that participation in these motifs provides stability (refs).
    Indeed, dynamic simulations suggest that participation in these motifs can provide greater persistence after a disturbance~\citep{Cirtwill2022Oikos}.
    The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
    Importantly, motif participation incorporates information on the positions of focal species' interaction partners, as well as potential indirect interactions between species.
    Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
    Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
    This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.

    Here, we focus on extinction risk after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
    The basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
    It is therefore crucial to understand how bottom-up effects of disturbance to basal resources affect the rest of the food web.
    As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.

    %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
    We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs (i.e., are more likely to retain all three species in the motif) when modelled in isolation~\citep{Borrelli2015a}.
    These four motifs also have clear ecological interpretations.
    The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
    The dynamics of three-species chains in isolation are well-understood from Lotka-Volterra models, and stable equilibria with all three species persisting can often be identified~\citep{Borrelli2015a,Hsu2015}.
    In general, food chains are more likely to be stable when predator-prey body-mass ratios are smaller~\citep{Jonsson1998}.
    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}. 
    Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey. The motif indicates a generalist predator (which may be able persist following reduction or loss of one prey) but can capture indirect effects between prey that make the predator more likely to go extinct~\citep{Melian2002}.
    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator. 
    Omnivory is less stable in isolation than the other three motifs~\citep{Borrelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra model of omnivory~\citep{Hsu2015}.
    In whole food webs, frequent omnivory has been proposed to both increase and decrease overall stability in different conditions~\citep{McCann1997,Emmerson2004,Monteiro2016}.
    As most of the above studies either model motifs in isolation or examine overall effects on the whole food web, it remains unclear how a particular species' motif participation will affect its specific extinction risk.

@@

        \begin{figure}[hb!]
        \centering
        % \includegraphics[width=.9\textwidth]{figures/Figure_1.png}
        \caption{Conceptual figure of a toy food web. \textbf{a)} This food web contains examples of the four motifs: apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow). These motifs capture different ways in which disturbances to basal resources can propagate through the network. For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species. 
        Note that the nodes participate in more motifs than the ones highlighted and that a species' motif participation vector is the count of all motifs of each type in which the species appears. \textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than three times more frequently (see inserted table for counts of participation in all four motifs, giving the motif participation vector in parentheses). 
        Despite having equal degree and trophic level, species A may be at greater risk of going extinct than species B because of its dependence on specialist prey species. Two-thirds of the prey of species B are generalists and therefore more likely to survive a disturbance to basal resources. By including information about indirect interactions, motif profiles reflect this difference between species A and B while trophic level and degree do not.}
    \label{fig:concept}
    \end{figure}

@@

        \begin{figure}[hb!]
        \centering
        % \includegraphics[width=0.85\textwidth]{figures/persistence_motif_participation.eps}
        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis). The effect of participating in each motif is based on the fixed effects in Equation~\ref{propreq}. The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance). 95\% confidence intervals for each line are shown in grey. Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
    \label{fig:prop_lmer_all}
    \end{figure}

    \begin{figure}[hb!]
    \centering
        % \includegraphics[width=\textwidth]{figures/Fig4.pdf}
        \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right). Rows show various levels of connectance. The dotted, vertical lines indicate zero on the x-axis. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence. The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend). }
        \label{fig:density_prop}
    \end{figure}

    \begin{figure}[ht!]
        \centering
        % \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
        \caption{The probabilities of persistence for a particular consumer (\textbf{A, B}) and the average probability of persistence across all consumers (\textbf{C-E}) varied with other elements of network structure as well as with motif participation.
        \textbf{A} Persistence of consumers decreased with increasing STL at all levels of disturbance.
        \textbf{B} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
        \textbf{C,D} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance. There was no significant relationship between mean persistence and network size.
        \textbf{E} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs. Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.}
        \label{fig:lm_CS}
    \end{figure}

    \begin{figure}[hb!]
        \centering
        % \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
        \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence. \textbf{A)} The proportion of omnivory increased with increasing degree while all other proportions decreased. \textbf{B)} The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased. \textbf{C-D)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition, where the interaction was not significant). To cover the largest range of slopes, we show relationships over connectance for the smallest networks (\textbf{C}) and over network size for the most-connected networks (\textbf{D}). See also Fig. SX. }
        \label{fig:motifs_vs_TL_and_deg}
    \end{figure}

>>>>

% Now 286 words. Max 350. 
    \indent{1. Loss of species in food webs can set in motion a cascade of additional, secondary,  extinctions.A species' position in a food web (e.g., its trophic level or number of interactions) is known to affect its ability to persist following disturbance. Theses simple measures, however, offer only a coarse description of how species fit into their community. Therefore one would expect that more detailed structure measures, such as a species' participation in three-species motifs (meso-scale structures which provide information on a species' direct and indirect interactions) will also be related to its probability of persistence.\\
    2. Disturbances, such as species extinctions, affecting the basal resources have particularly strong effects on the rest of the food web. However, how the disturbance branch out and affect consumer persistence is depending on the structural pattern of species interactions in several steps. The magnitude, e.g. proportion of basal resources lost, will likely also affect the outcome.\\     
    3. Here, we analyse whether a consumer's risk of going secondary extinct after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.\\
    4. We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.\\
    5. However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.\\
    6. Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied}

@@

    The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional (secondary) extinctions. 
    A species' risk of secondary extinction depends on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
    For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{Allesina2012,binzer2011susceptibility, Eklof2006}.
    These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources if a prey species is lost.
    The relationships between these relatively simple measures of a species' position in a food web suggest that more detailed measures, such as a species' participation in \emph{meso-scale motifs} will also be related to its persistence~\citep{Cirtwill2015a,Cirtwill2018FoodWebs,Cirtwill2022Oikos}.

    Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
    Some motifs are over-represented in stable food webs, suggesting that these motifs provides stability~\citep{Borrelli2015a}. This further suggest that the more frequently species are participating in these motifs, the more stable the food web would be.
    % Indeed, dynamic simulations show that participation in some motifs can provide greater persistence after the loss of a single species~\citep{Cirtwill2022Oikos}, suggesting that motif participation may be related to persistence after other types of disturbance.
    The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
    Importantly, motif participation incorporates information on the positions of focal species' interaction partners, as well as potential indirect interactions between species.
    Two species might have the same trophic level and the same number of prey, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (for an example see Fig.~\ref{fig:concept}).
    Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
    This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.

    %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
    We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs (i.e., are more likely to retain all three species in the motif) when modelled in isolation~\citep{Borrelli2015a}.
    These four motifs also have clear ecological interpretations.
    The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
    The dynamics of three-species chains in isolation are well-understood from Lotka-Volterra models, and stable equilibria with all three species persisting can often be identified~\citep{Borrelli2015a,Hsu2015}.
    In general, food chains are more likely to be stable when predator-prey body-mass ratios are smaller~\citep{Jonsson1998}.
    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}. 
   ( Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.)
    The apparent competition motif indicates a predator which consumes two resources and, by being somewhat generalist, may be able to persist following the loss or reduction of one prey.
    However, there may also be indirect effects between the linked prey that make the predator more likely to go extinct~\citep{Melian2002}.
    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes  the potential for a plant to have both direct and indirect effects on the predator. 
    Omnivory is less stable in isolation than the other three motifs~\citep{Borrelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra models of omnivory~\citep{Hsu2015}.
    In whole food webs, frequent omnivory has been proposed to both increase and decrease overall stability in different conditions~\citep{McCann1997,Emmerson2004,Monteiro2016}.
    As most of the above studies either model motifs in isolation or examine overall effects on the whole food web, it remains unclear how a particular species' motif participation will affect its specific extinction risk.

    Previous work using dynamical simulations has shown that participation in some motifs is associated with longer persistence after the loss of a single species in the food web~\citep{Cirtwill2022Oikos}, but the relationship between motif participation and persistence in the face of other disturbances remains unclear.
   As the basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
    It is therefore crucial to understand how bottom-up effects from disturbance to basal resources affect the rest of the food web.  Here, we are interested in understanding if motif participation can provide additional information on how extinction risk of consumer species are affected by the loss of basal resources, i.e., the effects of bottom-up disturbances.
    As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.

@@

        \begin{figure}[hb!]
        \centering
        \includegraphics[width=.9\textwidth]{figures/Figure_1.eps}
        \caption{Conceptual figure of a toy food web. \textbf{a)} This food web contains examples of the four motifs: apparent competition (orange), three-species chain (purple), omnivory (red), and direct competition (blue). These motifs capture different ways in which disturbances to basal resources can propagate through the network. For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species. 
        Note that the nodes participate in more motifs than the ones highlighted and that a species' motif participation vector is the count of all motifs of each type in which the species appears. \textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only three chains while species B (dark grey node with its direct consumption links colored dark grey) participates in seven (see inserted table for counts of participation in all four motifs; species A has the motif participation vector [3, 3, 0, 1] while species B has the motif participation vector [3, 7, 0, 1]).
        Despite having equal degree and trophic level, species A may be at greater risk of going extinct than species B because of its dependence on specialist prey species. Two-thirds of the prey of species B are generalists and therefore more likely to survive a disturbance to basal resources. By including information about indirect interactions, motif profiles reflect this difference between species A and B while trophic level and degree do not.}
    \label{fig:concept}
    \end{figure}

@@

        \begin{figure}[hb!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_motif_participation.eps}
        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on probability of persistence following disturbance to basal resources (y-axis). The effect of participating in each motif is based on the Equation~\ref{propreq}. The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance). 95\% confidence intervals for each line are shown in lighter shading. Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
    \label{fig:prop_lmer_all}
    \end{figure}

    \begin{figure}[hb!]
    \centering
        \includegraphics[width=\textwidth]{figures/Fig4.pdf}
        \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. Columns show the result for different motifs, rows show results for low, medium, and high connectance, and colour indicates the strength of disturbance to basal resources ($\pi_{disturbed}$; probability of extinction for a basal resource). The dotted, vertical lines indicate zero on the x-axis. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence.}
        \label{fig:density_prop}
    \end{figure}

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
        \caption{The probabilities of persistence for a particular consumer (\textbf{A, B}) and the average probability of persistence across all consumers (\textbf{C-E}) varied with other elements of network structure as well as with motif participation.
        \textbf{A} Persistence of consumers decreased with increasing STL at all levels of disturbance.
        \textbf{B} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
        \textbf{C,D} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance. There was no significant relationship between mean persistence and network size.
        \textbf{E} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs. Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.}
        \label{fig:lm_CS}
    \end{figure}

    \begin{figure}[hb!]
        \centering
        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
        \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence. \textbf{A)} The proportion of omnivory increased with increasing degree while all other proportions decreased. \textbf{B)} The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased. \textbf{C-D)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition, where the interaction was not significant). To cover the largest range of slopes, we show relationships over connectance for the smallest networks (\textbf{C}) and over network size for the most-connected networks (\textbf{D}). See also Fig. S6.}
        \label{fig:motifs_vs_TL_and_deg}
    \end{figure}
